The megamodeling language \ueber

### 2013-11-28 docs/ueber/paper.tex
Intermediate commit; ueber is broken right now.

\usepackage{tikz}
\usepackage{hyperref}
\usepackage{verbatim}
\usepackage{comment}
\usepackage{boxedminipage}
\usepackage{amssymb}
\usepackage{listings,lsthaskell,lstprolog,lstegl,lstesl,lstfsml,lstjava,lstdot}

@@

\newcommand{\wikipediafn}[2]{\footnote{\url{#1}---last visited #2.}}
\newcommand{\ooo}[1]{\textsf{101#1}}
\newcommand{\ueber}{\textsc{ueber}}

@@

\begin{abstract}
  We describe the megamodeling language \ueber, as it is used for
  modeling the linguistic architecture in the \slepro{} project; see
  \url{https://github.com/slebok/slepro}.

@@

\section{???}

@@

%\nocite{FreeDictionary}
%\setlength{\bibsep}{4pt}
%\bibliographystyle{elsarticle-harv}
%\bibliographystyle{elsarticle-num}
%\bibliography{paper}

>>>>

\usepackage{tikz}
\usepackage{hyperref}
\usepackage{verbatim}
\usepackage{comment}
\usepackage{boxedminipage}
\usepackage{amssymb}
\usepackage{listings,lsthaskell,lstprolog,lstegl,lstesl,lstfsml,lstjava,lstdot,lsttext,lstueber}

@@

\newcommand{\wikipediafn}[2]{\footnote{\url{#1}---last visited #2.}}
\newcommand{\ooo}[1]{\textsf{101#1}}
\newcommand{\ueber}{\textsf{ueber}}
\newcommand{\ppl}{\textsf{ppl}}
\newcommand{\esl}{\textsf{esl}}

@@

\begin{abstract}
  We describe the megamodeling language \ueber, as it is used for
  modeling the linguistic architecture of the Prolog-based \slepro{}
  project; see \url{https://github.com/slebok/slepro}.

@@

\section{Introducing \ueber}

The notion of megamodel has been defined and used in different ways in
software engineering. We favor here the view that a megamodel models
the architecture of a software system with particular focus on the
involved languages, technologies, and general software
concepts~\citep{FavreLV12}

In this document, we present a particular megamodeling language,
\ueber, which is used for megamodeling the Prolog-based \slepro{}
project, which is essentially a repository of language modeling
artifacts: parsers, translators, semantics definitions, etc.

In the following example, we demonstrate \ueber{} in a way as if we
were testing a language processor. Megamodeling is in no way focused
or restricted to software testing, but it happens that the particular
setup of the \slepro{} repository lets megamodeling be useful for
testing.

\codefigure{%
A \ppl{} expresion composing text literals vertically}{%
languages/ppl/test/vbox.ppl}{%
prolog}

\codefigure{%
The text rendered for the \ppl{} expression of \autoref{languages/ppl/test/vbox.ppl}}{%
languages/ppl/test/vbox.txt}{%
text}

We are concerned with a simple language \ppl---the Pretty Printing
Language.\footnote{\url{https://github.com/slebok/slepro/tree/master/languages/ppl}}
This language offers a number of simple composition combinators for
composing text vertically and horizontally. Consider the \ppl{}
expression of \autoref{languages/ppl/test/vbox.ppl}. \ppl{} comes with
a simple evaluation semantics that essentially renders the \ppl{}
expressions; see \autoref{languages/ppl/test/vbox.txt}.

\codefigure{%
Evaluation rule for vertical boxes}{%
languages/ppl/eval/vbox.pro}{%
prolog}

\codefigure{%
Final rendering step: turn lists of lines into a single string}{%
languages/ppl/to-string.pro}{%
prolog}

\autoref{languages/ppl/eval/vbox.pro} shows one rule of the evaluator
for \ppl{} terms---the rule for vertical
composition. \autoref{languages/ppl/to-string.pro} shows the trivial
conversion from lists of lines to a single string. This is a
straigthforward Prolog-based implementation of the pretty printing
combinators. The details do not matter much to the megamodeling
discussion that follows.

\codefigure{%
Syntax of \ppl{} terms}{%
languages/ppl/as.esl}{%
esl}

The syntax of all \ppl{} terms is defined in
\autoref{languages/ppl/as.esl}. We use a simple language, \esl, for
term algebras or algebraic data types, which we do not discuss here in
detail. The notation should be straightforward; we declare a number of
different symbols (`constructors') for different kinds of \ppl{}
combinators. The sample term of \autoref{languages/ppl/test/vbox.ppl}
conforms to the signature.

\codefigure{%
A megamodel for testing \ueber}{%
languages/ppl/test/.ueber}{%
ueber}

@@

%\setlength{\bibsep}{4pt}
\bibliographystyle{elsarticle-harv}
%\bibliographystyle{elsarticle-num}
\bibliography{paper}
